---
title: "Statistic Competition"
author: "Joel Chua"
date: "2025-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

working with data from UNSDG?
Motivation: Tracking progress of ASEAN countries on 17 UNDG
(likely will just pick the relevant ones)
Silver tide across Asian countries??
Alternative: Singapore data? on aging/demographic changes

#elderly's living arrangement in Singapore 
```{r}
library(readxl)
rawData_elderly <- read_excel("C:/Users/Joel Chua/OneDrive - National University of Singapore/Desktop/Poster competition/outputSingstat1.xlsx",sheet = "T2",range = "A11:P47")
```

#Pivot longer first
```{r}
library(tidyverse)
longData_elderly <- rawData_elderly %>%
  pivot_longer(cols = `2024`:`2010`,
              names_to = "year",
              values_to = "count") %>%
  rename(Category = `Data Series`)
```

#filling in the age categories
```{r}
longData_elderly <- longData_elderly %>%
  mutate(
    age_group = if_else(
      grepl("^[0-9]{2}.*Years", Category), Category, NA_character_)
  ) %>%
  fill(age_group) 
```

#Converting counts to proportions (within each year)
```{r}
df_long <- longData_elderly  %>%
  filter(!grepl("Years", Category)) %>%   # keep only living arrangement rows
  rename(living_arrangement = Category)

df_prop <- df_long %>%
  group_by(age_group, year) %>%
  mutate(
    total_age_group = sum(count),
    proportion = count / total_age_group
  ) %>%
  ungroup()
```


```{r}
library(ggplot2)
library(tidyverse)
df_prop_2024 <- df_prop %>% filter(year == 2024) %>% filter(!is.na(age_group))

library(ggplot2)

ggplot(df_prop, aes(x = age_group, y = proportion, fill = living_arrangement)) +
  geom_bar(stat = "identity", position = "fill", width = 0.6) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Proportion of Elderly by Living Arrangement for each Age Category (2024)",
    x = "Age Group",
    y = "Proportion"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#enhanced bar graphs for 2024
```{r}
library(ggplot2)
library(dplyr)

df_prop_2024 %>%
  ggplot(aes(x = age_group,
             y = proportion,
             fill = living_arrangement)) +
  geom_bar(stat = "identity", position = "fill", color = "white", linewidth = 0.2) +
  scale_y_continuous(labels = scales::percent_format()) +
  
  # Optional: reorder living arrangements (so similar types group visually)
  scale_fill_brewer(palette = "Set2") +
  
  labs(
    title = "Proportion of Elderly by Living Arrangement in 2024",
    subtitle = "Across Age Groups in Singapore",
    x = "Age Group",
    y = "Proportion",
    fill = "Living Arrangement"
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 13),
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right",
    panel.grid.minor = element_blank()
  )

```

```{r}
df_prop_2014 <- df_prop %>% filter(year == 2014) %>% filter(!is.na(age_group))

library(ggplot2)
library(dplyr)

df_prop_2014 %>%
  ggplot(aes(x = age_group,
             y = proportion,
             fill = living_arrangement)) +
  geom_bar(stat = "identity", position = "fill", color = "white", linewidth = 0.2) +
  scale_y_continuous(labels = scales::percent_format()) +
  
  # Optional: reorder living arrangements (so similar types group visually)
  scale_fill_brewer(palette = "Set2") +
  
  labs(
    title = "Proportion of Elderly by Living Arrangement in 2014",
    subtitle = "Across Age Groups in Singapore",
    x = "Age Group",
    y = "Proportion",
    fill = "Living Arrangement"
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 13),
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "right",
    panel.grid.minor = element_blank()
  )
```


# data on hospital admission rates 
```{r}
library(readxl)
rawData_eld2 <- read_excel("C:/Users/Joel Chua/OneDrive - National University of Singapore/Desktop/Poster competition/outputFile.xlsx",sheet = "T1",range = "A10:V20")
```

```{r}
library(tidyverse)
library(dplyr)
year_cols <- grep("^20", names(rawData_eld2), value = TRUE)

rawData_eld2 <- rawData_eld2 %>%
  mutate(across(all_of(year_cols), ~ as.numeric(.)))

longData_2 <- rawData_eld2 %>%
  pivot_longer(cols = `2025`:`2005`,
              names_to = "Year",
              values_to = "Value") %>%
  rename(Indicator = `Data Series`)
```

#data on Proportion Of Elderly Residents (65 Years & Over) In Labour Force

```{r}
library(tidyverse)
library(dplyr)
longData_laborForce <- longData_2 %>% 
  slice(86:105)


ggplot(longData_laborForce, aes(x = Year, y = Value,group = 1)) +
  geom_line() +
  labs(
    title = "Proportion Of Elderly Residents (65 Years & Over) In Labour Force (Per Cent)",
    x = "Year",
    y = "% of Elderly in Labour Force"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 10, face = "bold", hjust = 0.5)
  )
```

# Sex ratio graph
```{r}
library(tidyverse)
library(dplyr)
longData_sexRatio <- longData_2 %>% 
  slice(65:84)


ggplot(longData_laborForce, aes(x = Year, y = Value, group = 1)) +
  geom_col(fill = "skyblue") +
  labs(
    title = "Sex Ratio Of Elderly Residents (65 Years & Over)",
    x = "Year",
    y = "Males Per Thousand Females"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 10, face = "bold", hjust = 0.5)
  )
```


#life expectancy at 65 y.o
```{r}
library(tidyverse)
library(dplyr)
longData_lifeExpectancy <- longData_2 %>% 
  slice(128:147)


ggplot(longData_lifeExpectancy, aes(x = Year, y = Value, group = 1)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Life Expectancy At Age 65 Years (Years)",
    x = "Year",
    y = "Life Expectancy At Age 65 Years (Years)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 10, face = "bold", hjust = 0.5)
  )
```

#death rate, heart disease rate and cancer rate per 1000
```{r}
healthData <- longData_2 %>% 
  slice(149:n()) %>% 
  filter(Year < 2025 & Year > 2014)

ggplot(healthData, aes(x = Year, y = Indicator, fill = Value)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightyellow", high = "red") +
  labs(
    title = "Health Outcomes Among Elderly Residents (2014–2024)",
    x = "Year",
    y = "Health Indicator",
    fill = "Rate"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", size = 18)
  )

```

